---
title: "HW7"
author: "Melchor Ronquillo"
date: "2022-12-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

8. This problem involves the OJ data set which is part of the ISLR2 package.

    (a) Create a training set containing a random sample of 800 observations, 
        and a test set containing the remaining observations.
```{r}
install.packages('ISLR2', repos = "http://cran.us.r-project.org")
library(ISLR2)

Juice <- OJ
Juice


set.seed(777)
randomJuice= Juice[sample(1:nrow(Juice)), ]
#randomJuice

nrow(randomJuice)
train_OJ <- randomJuice[1:800,]
test_OJ <- randomJuice[801:1070,]

#train_OJ
#test_OJ
```

    (b) Fit a support vector classifier to the training data using cost = 0.01, 
        with Purchase as the response and the other variables as predictors. 
        Use the summary() function to produce summary statistics, and describe 
        the results obtained.
```{r}
install.packages('e1071', repos = "http://cran.us.r-project.org")
library(e1071)

set.seed(778)
JuiceSVC <- svm(Purchase ~ ., data = train_OJ, kernel = "linear", cost = 0.01, scale = TRUE)
#plot(JuiceSVM, train_OJ)
summary(JuiceSVC)
```

          
        
    (c) What are the training and test error rates?
```{r}
train_OJ_pred = predict(JuiceSVC, train_OJ)
table(train_OJ$Purchase, train_OJ_pred)
train_error = (57+75) / (453+57+75+235)
train_error
```

        Train error rate:
          16% error
 
```{r}
test_OJ_pred = predict(JuiceSVC, test_OJ)
table(test_OJ$Purchase, test_OJ_pred)
test_error = (26+24) / (137+26+24+83)
test_error
```
        
        Test error rate:
          18.5%
        
        
    (d) Use the tune() function to select an optimal cost. Consider values in 
        the range 0.01 to 10.
```{r}
set.seed(779)
tuneJuice <- tune(svm, Purchase ~ ., data = Juice, kernel = "linear", 
                  ranges = list(cost = c(0.01, 0.05, 0.1, 0.5, 1, 5, 10))
                  )
summary(tuneJuice)
```

    (e) Compute the training and test error rates using this new value for cost.
```{r}
bestmod <- tuneJuice$best.model
summary(bestmod)

bestpred_OJ_SVC <- predict(bestmod, test_OJ)
table(test_OJ$Purchase, bestpred_OJ_SVC)
test_error = (23+23) / (140+23+23+84)
test_error
```

        Test Error: 17%
        
        
    (f) Repeat parts (b) through (e) using a support vector machine with a 
        radial kernel. Use the default value for gamma.
```{r}
#Create SVM
set.seed(780)
JuiceSVM <- svm(Purchase ~ ., data = train_OJ, kernel = "radial", 
                gamma = 1, cost = 0.01, scale = TRUE)
#plot(JuiceSVM, train_OJ)
summary(JuiceSVM)

#train and test errors
train_OJ_predSVM = predict(JuiceSVM, train_OJ)
table(train_OJ$Purchase, train_OJ_predSVM)

test_OJ_predSVM = predict(JuiceSVM, test_OJ)
table(test_OJ$Purchase, test_OJ_predSVM)

#Tune
set.seed(781)
tuneJuiceSVM <- tune(svm, Purchase ~ ., data = Juice, kernel = "radial", 
                  ranges = list(cost = c(0.01, 0.05, 0.1, 0.5, 1, 5, 10))
                  )

summary(tuneJuiceSVM)

bestmodSVM <- tuneJuiceSVM$best.model
summary(bestmodSVM)

bestpred_OJ_SVM <- predict(bestmodSVM, test_OJ)
table(test_OJ$Purchase, bestpred_OJ_SVM)
test_error = (14+26) / (149+14+26+81)
test_error
```

        Test Error: 14.8%
   
   
   
    (g) Repeat parts (b) through (e) using a support vector machine with a 
        polynomial kernel. Set degree = 2.
```{r}
set.seed(782)
JuiceSVCPoly <- svm(Purchase ~ ., data = train_OJ, kernel = "polynomial", 
                    cost = 0.01,
                    degree = 2,
                    scale = TRUE)
summary(JuiceSVCPoly)

#train and test errors
train_OJ_predPoly = predict(JuiceSVCPoly, train_OJ)
table(train_OJ$Purchase, train_OJ_predPoly)

test_OJ_predPoly = predict(JuiceSVCPoly, test_OJ)
table(test_OJ$Purchase, test_OJ_predPoly)

#tune
set.seed(779)
tuneJuicePoly <- tune(svm, Purchase ~ ., data = Juice, kernel = "polynomial",
                      degree = 2, scale = TRUE,
                      ranges = list(cost = c(0.01, 0.05, 0.1, 0.5, 1, 5, 10))
                  )
summary(tuneJuicePoly)

#best model
bestmodPoly <- tuneJuicePoly$best.model
summary(bestmodPoly)

bestpred_OJ_Poly <- predict(bestmodPoly, test_OJ)
table(test_OJ$Purchase, bestpred_OJ_Poly)
test_error = (15+25) / (148+15+25+82)
test_error
```

        Test Error: 14.8%
    
    (h) Overall, which approach seems to give the best results on this data?
    
        Overall, the approach witht he best results was both radial and polynomial, 
        as they both had the lowest test error rate of 14.8%
